\input{/Users/daniel/github/config/preamble.sty}%available at github.com/danimalabares/config
\input{/Users/daniel/github/config/thms-eng.sty}%available at github.com/danimalabares/config


\usepackage[style=authortitle-terse,backend=bibtex]{biblatex}
\addbibresource{~/github/config/bibliography.bib}

\setcounter{secnumdepth}{2}

\begin{document}

\begin{minipage}{\textwidth}
	\begin{minipage}{1\textwidth}
		K3 surfaces \hfill Daniel González Casanova Azuela
		
		{\small Prof. Misha Verbitsky\hfill\href{https://github.com/danimalabares/k3}{github.com/danimalabares/k3}}
	\end{minipage}
\end{minipage}\vspace{.2cm}\hrule

\vspace{10pt}
{\huge Home assignment 7: Moser isotopy lemma}

\begin{thing1}{Exercise 7.1}\leavevmode
Let $M$ be a compact manifold, and $V_0,V_1$ two smooth volume forms which satifsy  $ \int_{M}V_0=\int_{M}V_1$. Prove that there exists a diffeomorphism which satisfies $\Phi^*V_0=V_1$.
\end{thing1}
\begin{proof}[Solution]\leavevmode
First notice that the condition $\int_{M}V_0=\int_{M}V_1$ implies that $V_0$ and $V_1$ are cohomologous. This follows from the fact that $H^n(M)$ is a real one-dimensional vector space, so $[V_1]=a[V_0]$ for some real number  $a$. Then
\begin{align*}
	\alpha\int_{M}[V_0]=\int_{M}[V_1]&=\int_{M}[V_0]
\end{align*}
which means that $a=1$. This means that there exists a  ($n-1$)-form $\eta$ such that
\[V_1-V_0=d\eta.\]
Now define
\[V_t=V_0+t\left(V_1-V_0\right),\qquad t\in[0,1].\]

Notice that $V_t$ is a nowhere-vanishing top-form. This follows since $ V_0$ and $V_1$ are nowhere-vanishing and their integrals over $M$ coincide, so they are both always positive or always negative. Now if $V_t=0$ we would have $t=1/2$ but  $V_{1/2}=1/2(V_1+V_0)$.

To solve this exercise we will immitate the proof of Moser lemma given in lecture 16, which coincides with the proof in \cite{lee} and in this \href{https://math.stackexchange.com/questions/3528448/equal-volume-iff-there-is-a-diffeomorphism}{StackExchange} answer.

First notice that the contraction map
\begin{equation}\label{eq:1}
\begin{aligned}
	i_\bullet V_t: \mathfrak{X}(M) &\longrightarrow \Omega^{n-1}(M) \\
	X &\longmapsto i_X V_t
\end{aligned}
\end{equation}
is a bundle isomorphism for every $t$. The proof of this fact was found in yet another \href{https://math.stackexchange.com/questions/3403451/if-omega-is-a-volume-form-then-x-mapsto-i-x-omega-generates-all-n-1-fo}{StackExchange} post. First we notice that this map is an isomorphism at every point. Indeed, if $i_{X_1}V_t=i_{X_2}V_t$, then
\[V_t(X_1-X_2,\cdot,\ldots,\cdot)=0.\]
This is not possible since if $X_1-X_2\neq 0$, such vector can be completed to a basis where $V_t$ cannot be zero since it is a volume form. Further, $i_\bullet V_t$ is an isomorphism at every point since both $T_pM$ and $V^{n-1}(T_pM)$ have dimension $n$.

We must also make sure that this operation is smooth. This can be checked in a local chart $U$ where $V_t=gdx_1\ldots dx_n$. Choose any differential form $\alpha=\sum f_idx_1\ldots\widehat{dx_i}\ldots dx_n\in\Omega^{n-1}(U)$. Smoothness follows from the fact that the coordinate functions of its corresponding vector field are also smooth. Indeed, we claim that $i_XV_t=\alpha$ for
\[X=\sum (-1)^{i+1}\frac{f_i}{g}\partial_i,\]
whose coordinate functions are smooth.

This is most easily seen in the case $n=2$. By Leibniz rule of interior multiplication,
\begin{align*}
i_X(gdx_1\wedge dx_2)&=(i_Xgdx_1)\wedge dx_2-gdx_1\wedge i_Xdx_2\\
&=f_1 dx_2- f_2 dx_1=\alpha
\end{align*}
A similar computation confirms the general case.

Having established the isomorphism in \cref{eq:1}, we see that $-\eta \in \Omega^{n-1}(M)$ can be seen as $i_{X_t}V_t$ for some vector field $X_t$ for each  $t$. Now we will show that the pullback $\varphi^*_t \omega_t$ is constant as $t$ varies. The proof relies on the following result: 

\begin{thing4}{Proposition 22.15}[\cite{lee}]\leavevmode
	Let $\{X_t\}$ be an autonomous (time-dependent) vector field and choose for every $t$ a $k$-form $\omega_t \in \Omega^{k}(M)$. Then
	\[\frac{d}{dt}(\varphi^* _t \omega_t)=\varphi^*\left( \mathcal{L}_{X_t}\omega_t+\frac{d}{dt}\omega_t \right),\]
	where $\varphi_t$ is the isotopy corresponding to $\{X_t\}$, i.e. the family of flows associated to $\{X_t\}$.
\end{thing4}
So we have
\begin{align*}
\frac{d}{dt}\varphi^*_tV_t&=\varphi^*\left( \mathcal{L}_{X_t}V_t+\frac{d}{dt}V_t \right)\\
&=\varphi^*\left( di_{X_t}V_t+\cancelto{0}{i_{X_t}dV_t}+\frac{d}{dt}\left( V_0+t(V_1-V_0) \right)\right)\\
&=\varphi^*\left( di_{X_t}V_t+V_1-V_0 \right)\\
&=\varphi^*\left(   d(-\eta)+d\eta\right)=0
\end{align*}
Evaluating at $t=1$ we see that $\varphi_1$ is the desired isomorphism since
\[\varphi_1^*V_1=\varphi^*_0V_0=\operatorname{id}^*V_0=V_0.\]

\end{proof}
\begin{thing5}{Problem 7.2}\leavevmode
	Let $(M,I)$ be an almost complex manifold, and $\omega_0,\omega_1$ co-homologous symplectic forms which satisfy $\omega_i(x,Ix)>0$ for any non-zero tangent vetor $x$ (such forms are called \textit{\textbf{taming}}).
	\begin{enumerate}[label=\alph*.]
	\item Prove that there exists a diffeomorphism $\Phi$ which satisfies $\Phi^*\omega_0=\omega_1$.
	\item Prove that $|x|^2_{i}:=\omega_i(X,Ix)$ is a Hermitian metric on $M$. Prove that a diffeomorphism that satisfies $\Phi^*\omega_0=\omega_1$ defines an isometry
		\[(M,|x|^2_1)\longrightarrow(M,|x|^2_0)\]
		is $\Phi$ is compatible with $I$, that is, satisfies $d\Phi(Ix)=I(d\Phi(x))$ \item Find an example of $\omega_0,\omega_1\in\Lambda^2(M)$ such that a diffeomorphism compatible with $I$ and satisfying $\Phi^*\omega_0=\omega_1$ does not exist.
	\end{enumerate}
\end{thing5}

\begin{proof}[Solution]\leavevmode
\begin{enumerate}[label=\alph*.]
\item This is immediate from Moser's lemma defining $\omega_t:=t\omega_0+(1+t)\omega_1$.
\item According to Riemann Surfaces course, a hermitian form may be understood as a Riemannian metric $h$ such that $h(\cdot,\cdot)=h(I\cdot,I\cdot)$. This happens for $h(\cdot,\cdot)=\omega(\cdot,I\cdot)$ if we require $\omega(\cdot,\cdot)=\omega(I\cdot,I\cdot)$. Indeed, symmetry holds since
	\[\omega(x,Iy)=\omega(Ix,-y)=\omega(y,Ix)\]
and positive-definiteness is given as a hypothesis.
	

	The condition that $\omega(\cdot,I\cdot)$ is a riemannian metric is called \textit{\textbf{compatibility}} of  $\omega$ and $I$ in  \cite{das}. In this case we can also produce a hermitian metric in the sense of  \href{https://mathworld.wolfram.com/HermitianInnerProduct.html}{WolframMathWorld}, namely, a positive-definite symmetric sesquilinear form. Indeed, define $g(\cdot,\cdot):=\omega(\cdot,I\cdot)$, then the form
	\[h:=g+\sqrt{-1}\omega\]
satisfies the required properties as follows.
	\begin{enumerate}
	\item Additivity, $h(u_1+u_2,v)=h(u_1,v)+h(u_1,v)$, is immediate.
	\item Homegenity on the first argument, $h(\lambda u,v)=\lambda h(u,v)$, is also immediate.

			\item $h(u,v)=\overline{h(v,u)}$ is clear by anti-symmetry of $\Omega$:
			\begin{align*}
				h(u,v)& =g(u,v)+i\omega(u,v) =g(v,u)-i\omega(v,u) =\overline{h(v,u)}
			\end{align*}
	\item The property $h(u,\lambda v)=\bar{\lambda} h(u,v)$ follows easily from (b) and (c) since 
			\begin{align*}
				h(u,\lambda v)& =\overline{h(\lambda v, u)}=\bar{\lambda} \overline{h(v,u)} =\bar{\lambda} h(u,v)
			\end{align*}
	\item Positive-definiteness follows from positive-definiteness of $g$ and antisymmetry of $\omega$.
	\end{enumerate}
	\item The conditions $\Phi^*\omega_0=\omega_1$ and $\Phi_*(I\cdot)=I\Phi_*(\cdot)$ imply that
		\begin{equation}\label{eq:1}(\Phi^*\omega_0)(\cdot,I\cdot)=\omega_0(\Phi_*\cdot,\Phi_*I\cdot)=\omega_0\Phi_*\cdot,I\Phi_*\cdot)=\omega_1(\cdot,I\cdot)\end{equation}
	Whether the metric is taken to be $\omega_i(\cdot,I\cdot)$ or $h_i=g_i+\sqrt{-1}\omega_i$, the result follows. In the latter case only note that
	\begin{align*}\Phi^*h_0&=\Phi^*(g_0+\sqrt{-1}\omega_0)=\Phi^*g_0+\sqrt{-1}\Phi^*\omega_0\\&=\Phi^*\Big(\omega(\cdot,I\cdot)\Big)=g_1+\sqrt{-1}\omega_1=h_1\end{align*}
	by \cref{eq:1}.
	
\item I am intrigued to know the answer here.	
\end{enumerate}
\end{proof}

\begin{thing8}{Exercise 7.4}
Let $(M,V)$ be a connected manifold equipped with a volume form. Prove that the group of volume-preserving diffeomorphisms acts on $M$ transitively.
\end{thing8}

\begin{proof}[Solution]\leavevmode
I will sketch the proof of  \cite{boothby}, found in \href{https://mathoverflow.net/questions/61994/how-transitive-are-the-actions-of-symplectomorphism-groups}{MathOverflow}.

\iffalse The idea is to split a path joining any two points into small pieces, and find the volume-preserving diffeomorphisms locally.

\begin{enumerate}
\item \textit{\textbf{Strong local transitivity}} on $M$ means that for every point $p\in M$ and neighbourhood $U$ of $p$ there are neighbourhoods $V$ and $W$ of $p$ with $\overline{V}\subset W$ and $\overline{W} \subset U$, $\overline{W}$ compact, and for any $q \in V$ there is a volume-preserving isotopy $\Phi_t$ of the identity map $\Phi_0$ to $\Phi_1$ such that $\Phi_1(p)=q$ and for all  $t$, $\Phi_t$ leaves fixed every point outside $\overline{W}$.
\item If $M$ is strongly locally transitive then it is transitive by volume-preserving diffeomorphisms. \textit{Sketch of proof:} let $\gamma$ be a path from $p$ to $q$. For every $x \in \gamma$ choose $0<\delta'_x<\delta''_x<\varepsilon$  such that the $\delta'$ and $\delta''$ neighbourhoods of $x$ satisfy the conditions of strong local transitivity. Since $C_1$ is compact, there is a finite collection of points $p_1=a_1,a_2,\ldots,a_r=q_1$ with volume-preserving diffeomorphisms mapping each point to the next. The composition of all such diffeomorphisms yields the desired volume-preserving map.
\item It remains to show that any connected manifold with a volume form is strongly locally transitive. Consider a path $\gamma$ from $p$ to $q$.
\fi
	
	%(such a neighbourhood must be chosen correctly to satisfy the conditions of strong locally transivitity, namely there must be two subneighbourhoods $V$ and  $W$ as above).
We wish to define a vector field along $\gamma$ whose flow $\Phi_t$ gives the desired volume preserving diffeomorphism at $t=1$. Consider a diffeomorphism from $I\times B^{n-1}$ to a neighbourhood $U$ of $\gamma$ (this diffeomorphism is constructed from geodesic segments introducing a metric on $M$). The volume form on $U$ may be expressed as \[V=fdt \wedge dx_1 \wedge\ldots\wedge dx_{n-1}\] for a positive function $f$. Now consider the vector field
\[X'=\frac{1}{f}\frac{\partial }{\partial t}.\]
$X'$ may be extended to a vector field $X$ in all of $M$ using a partition of unity. It is immediate that the flow of this vector field takes $p$ to $q$.

To show this flow is volume-preserving recall Cartan's formula that $\mathcal{L}_XV=di_{X}V+\cancelto{0}{i_XdV}=di_XV$ since $V$ is a top-form. Thus, showing volume invariance ammounts to showing $d_XV=0$, which holds since
\[i_{X'}fdt\wedge dx_1 \wedge \ldots\wedge dx_{n-1}=dx_1\wedge \ldots \wedge dx_{n-1}\]
is closed.
%\end{enumerate}
\end{proof}

\begin{thing5}{Exercise 7.5}\leavevmode
	Prove that the group of symplectomorphisms of $(M,\omega)$ acts transitively on $M$, for any connected symplectic manifold $(M,\omega)$.
\end{thing5}

\begin{proof}[Solution]\leavevmode
This case is similar but requires a little more work: we intend to use Darboux coordinate charts, which can only be defined locally. We will split a path joining any two points into small pieces, find the symplectomorphisms locally and the compose them.
\begin{enumerate}
\item \textit{\textbf{Strong local transitivity}} on $M$ means that for every point $p\in M$ and neighbourhood $U$ of $p$ there are neighbourhoods $V$ and $W$ of $p$ with $\overline{V}\subset W$ and $\overline{W} \subset U$, $\overline{W}$ compact, and for any $q \in V$ there is a symplectomorphism isotopy $\Phi_t$ of the identity map $\Phi_0$ to $\Phi_1$ such that $\Phi_1(p)=q$ and for all  $t$, $\Phi_t$ leaves fixed every point outside $\overline{W}$.
\item If $M$ is strongly locally transitive then it is transitive by symplectomorphisms. \textit{Sketch of proof:} let $\gamma$ be a path from $p$ to $q$. For every $x \in \gamma$ choose $0<\delta'_x<\delta''_x<\varepsilon$  such that the $\delta'$ and $\delta''$ neighbourhoods of $x$ satisfy the conditions of strong local transitivity. Since $\gamma_1$ is compact, there is a finite collection of points $p_1=a_1,a_2,\ldots,a_r=q_1$ with volume-preserving diffeomorphisms mapping each point to the next. The composition of all such diffeomorphisms yields the desired volume-preserving map.
\item It remains to show that any connected symplectic manifold is strongly locally transitive. Let $U$ be a Darboux chart of $p$ and take two neighbourhoods $V'$ and $W'$ as in the definition of strong local transitivity. For any  $y \in V'$ we can define the vector field $X$ in $\mathbb{R}^{2n}$ of vectors parallel to $yx_0$, where $x_0$ are the coordinates of $p$. The flow of this vector field takes takes $x_0$ to $y$ at $t=1$.

To show that this vector field is $\omega$-invariant we use again Cartan's formula to see it's enough to show that $di_X\omega=0$. Now if $\dim M=2m$ we have
\[i_X\omega=\sum_{i=1}^m dx^i\wedge dx^{i+m}\left(X,\cdot \right){\color{8}=0}\]
({\color{8}I'm not sure why does it vanish…})
\end{enumerate}


\end{proof}

\vspace{1em}

{\Large Extra exercise: infinite product}\vspace{1em}

\begin{thing4}{Exercise 8}[\cite{zo1}, 3.2.5, p. 148]\leavevmode
An inifinite product $\prod_{k=1}^\infty e_k$ is said to converge if the sequence $\Pi_n=\prod_{k=1}^n e_k$ has a finite \textbf{nonzero} limit $\Pi$. We then set $\Pi=\prod_{k=1}^\infty e_k$. Show that
\begin{enumerate}[label=(\alph*)]
	\item[(b)]  if $\forall n \in N (e_n>0)$, then the infinite product converges $\prod_{n=1}^\infty$ converges if and only if the series  $\sum_{n=1}^\infty \operatorname{log}e_n$ converges.
\end{enumerate}
\end{thing4}

\begin{remark}\leavevmode
The proof is straightforward using elementary properties of the logarithm ($\operatorname{log}(ab)=\operatorname{log}a+\operatorname{log}b$) and the exponential ($\operatorname{exp}(a+b)=\operatorname{exp}a+\operatorname{exp}b$), as well as their continuity. The caveat is that if the product converges to zero, taking logarithms of the partial products gives a series diverging to $-\infty$ (we say that the product \textit{\textbf{diverges to zero}}). The following exercise is an example of this.
\end{remark}

\begin{claim}\leavevmode
\begin{align*}
			\prod_{n\geq 0}\left|  -1+\frac{1}{n+\frac{1}{2}}\right|=0.
				\end{align*}
\end{claim}			
\begin{proof}\leavevmode
By taking exponent of the partial sums of the following series, we see it's enough to show that
\begin{equation}\label{eq:log}\sum_{n \geq 0}\operatorname{log}\left|  -1+\frac{1}{n+\frac{1}{2}}\right|=-\infty.
\end{equation}
To show this first notice that $\left|  -1+\frac{1}{n+1/2}\right|$ is just  $1-\frac{1}{n+1/2}$. We can also quickly notice that $\operatorname{log}\left(1-\frac{1}{n+1/2}\right)$ is a sequence of negative numbers convering to 0.

(It was not immediate to me why the this series should diverge or converge. The following argument was provided by ChatGPT.)

We can prove the series diverges if we find a divergent series that bounds it from above. Such series is $\sum\frac{-1}{n+1/2}$. Indeed, it turns out that for every number $0\leq x<1$,
\[\operatorname{log}(1-x) \leq -x.\]
To see why, define the function $f(x)=\operatorname{log}(1-x)+x$ and differentiate to find
\[\frac{d}{dx}\operatorname{log}(1-x)+x=-\frac{1}{1-x}+1=\frac{1}{1-x}+\frac{1-x}{1-x}=\frac{-x}{1-x}\leq 0,\]
and since also $f(0)=0$, we have that
\[\operatorname{log}(1-x)+x\leq 0,\qquad 0\leq x<1.\]
Finally just recall that $\sum_{n\geq 0}\frac{1}{n}$ diverges (and behaves like $\sum \frac{1}{n+1/2}$ for lage $n$). This confirms \cref{eq:log}.




\end{proof}

\printbibliography




\end{document}
